---
title: "CLass Demo 10-5"
author: "Wells"
date: "10/5/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```


## Changing Classification Threshold in Log-Reg Models

Let's use the iris data set

```{r}
data(iris)
```

The `Species` variable is categorical and will serve as our response. Since log-reg is only appropriate for binary categorical variables, we create a new response `setosa` based on whether a given iris setosa or not.

```{r}
iris1 <- iris %>% mutate(setosa = ifelse(Species == "setosa", 1, 0))
```

Now we create the logistic regression model, with `setosa` as a function of `Sepal.Width`.

```{r}
log_reg <- glm(setosa ~ Sepal.Width , data = iris1, family = "binomial")
```

We create probability predictions from the model (note that the result would be log-odds if we omitted `type = "response"`)

```{r}
probs<-predict(log_reg, iris1, type = "response")
```

Finally, we create our actual predictions using an `ifelse` statement. The classification threshhold is determined by the value we set on the right side of the `>=` inequality. Here, for example, we set our threshold at 0.253.

```{r}
predicts<-ifelse( probs >= .253, 1, 0)
predicts
```

